Surrogate Regret Bounds for Polyhedral Losses

### 2021-05-26 rates-neurips-21/rates.tex
pretty up section 5 a bit

Our main upper bound for polyhedral losses, Theorem \ref{thm:main-upper}, is mostly nonconstructive: it says polyhedral surrogates have some $\alpha$ such that $R_{\ell}(\psi \circ h; \D) \leq \alpha \cdot R_L(h;\D)$.
In this section, we make the result constructive by unpacking the implications of consistency (or more precisely, calibration).

We will give a bound on the constant of the form $\alpha \leq \alpha_{\ell} \cdot \alpha_L \cdot \alpha_{\psi}$, i.e. one component each relating to the structure of the target loss, surrogate loss, and link.
First, we will derive $\alpha_L$ using the theory of \emph{Hoffman constants} from linear optimization.
Then, we will show that \emph{any} link for a consistent polyhedral loss must satisfy a condition called $\epsilon$-separation introduced in \cite{finocchiario2020}.
We will set $\alpha_{\psi} = \frac{1}{\epsilon}$.
With $\alpha_{\ell}$ simply an upper bound on the target loss, we will finally prove $\alpha \leq \alpha_{\ell} \cdot \alpha_L \cdot \alpha_{\psi}$.

@@

\cite{finocchiaro2020} showed that, if $L$ indirectly elicits $\ell$, then there exists an $\epsilon$-separated link $\psi$ such that $(L,\psi)$ is consistent for $L$.
Here, we give a strengthening and converse that we believe is of independent interest: not only that link, but \emph{any} link making $L$ consistent is $\epsilon$-separated for some $\epsilon > 0$.
Then, we show that this separation implies linear regret transfers (with a concrete constant).

\begin{lemma}\label{lem:calibrated-eps-sep}
  Let polyhedral surrogate $L:\reals^d \to \reals^\Y_+$, discrete loss $\ell:\R\to\reals^\Y_+$, and link $\psi:\reals^d\to\R$ be given such that $(L,\psi)$ is calibrated with respect to $\ell$.
  Then there exists $\epsilon>0$ such that $\psi$ is $\epsilon$-separated with respect to   $\Gamma \defeq \prop{L}$ and $\gamma \defeq \prop{\ell}$.
\end{lemma}
\begin{proof}
  Suppose that $\psi$ is not $\epsilon$-separated for any $\epsilon>0$.
  Then letting $\epsilon_i \defeq 1/i$ we have sequences $\{p_i\}_i \subset \simplex$ and  $\{u_i\}_i \subset \reals^d$ such that for all $i\in\mathbb N$ we have both $\psi(u_i) \notin \gamma(p_i)$ and $d_\infty(u_i,\Gamma(p_i)) \leq \epsilon_i$.
  First, observe that there are only finitely many values for $\gamma(p_i)$ and $\Gamma(p_i)$, as $\R$ is finite and $L$ is polyhedral.
  \raft{Using a result from the polyhedral paper here, that $\Gamma$ only takes on finitely many values}
  Thus, there must be some $p\in\simplex$ and some infinite subsequence indexed by $j\in J \subseteq \mathbb N$ where
  for all $j\in J$, we have $\psi(u_j) \notin \gamma(p)$ and $\Gamma(p_j) = \Gamma(p)$.

@@

\subsection{Combining the loss and link}
We can now follow the general proof strategy of the main upper bound, but constructively.
We will need one more ingredient, an upper bound on the regret of $\ell$.
\begin{definition}
  Given a discrete target loss $\ell: \R \to \reals_+^{\Y}$, define $C_{\ell} = \max_{r,r' \in \R, y \in \Y} \ell(r)_y - \ell(r')_y$.
\end{definition}
Observe that $C_{\ell} = \max_{r,p} R_{\ell}(r,p)$.
\bo{needs justification?}
\bo{We could have $C_{\ell} = 0$ but then everything is optimal and the regret transfer is zero.}

@@

\begin{theorem}[Constructive linear transfer] \label{thm:separated-constant}
  Let $\ell: \R \to \reals_+^{\Y}$ be a discrete target loss, $L: \reals^d \to \reals_+^{\Y}$ be a polyhedral surrogate loss, and $\psi: \reals^d \to \R$ a link function.
  If $(L,\psi)$ indirectly elicit $\ell$ and $\psi$ is $\epsilon$-separated, then
    \[ (\forall h,\D) \quad R_{\ell}(\psi \circ h ; \D) \leq \frac{C_{\ell} H_L}{\epsilon} R_L(h ; \D) ~. \]
\end{theorem}
The proof closely mirrors the proof of the nonconstructive upper bound, Theorem \ref{thm:main-upper}.
\begin{proof}
  By Lemma \ref{lemma:separated-constant-p}, for each $p \in \mathcal{Q}$, $R_{\ell}(\psi(u),p) \leq \frac{C_{\ell} H_L}{\epsilon} R_L(u,p)$ for all $u$.
  Now consider a general $p$, which is in some full-dimensional polytope level set $\Gamma_u$.
  Write $p = \sum_{q \in \mathcal{Q}_u} \beta(q) q$ for some probability distribution $\beta$, where $\mathcal{Q}_u$ is the set of vertices of $\Gamma_u$.
  By Lemma \ref{lemma:linear-on-levelset}, $R_L$ and $R_{\ell}$ are linear in $p$ on $\Gamma_u$, so for any $u'$,
  \begin{align*}
    R_{\ell}(\psi(u'),p)
    &=    \sum_{q \in \mathcal{Q}_u} \beta(q) R_{\ell}(\psi(u'), q)  \\
    &\leq \sum_{q \in \mathcal{Q}_u} \beta(q) \frac{C_{\ell} H_{L,p}}{\epsilon} R_L(u', q)  \\
    &\leq \frac{C_{\ell} H_L}{\epsilon} \sum_{q \in \mathcal{Q}_u} \beta(q) R_L(u', q)  \\
    &\leq \frac{C_{\ell} H_L}{\epsilon} R_L(u', p) .
  \end{align*}
  This conditional regret transfer implies a full regret transfer with the same constant by Observation \ref{obs:transfer}.
\end{proof}

@@

\begin{theorem} \label{thm:tfae}
  Let $\ell: \R \to \reals_+^{\Y}$ be a discrete target loss, $L: \reals^d \to \reals_+^{\Y}$ be a polyhedral surrogate loss, and $\psi: \reals^d \to \R$ a link function.
  The following are equivalent:
  \begin{enumerate}
    \item $(L,\psi)$ are consistent for $\ell$.
    \item $(L,\psi)$ indirectly elicit $\ell$ and $\psi$ is $\epsilon$-separated for some $\epsilon > 0$.
    \item $(L,\psi)$ guarantee a linear regret transfer for $\ell$.
  \end{enumerate}
  Furthermore, if any of the above hold, then in particular $(L,\psi)$ guarantee the regret transfer $\zeta(t) = \left(\frac{C_{\ell} H_L}{\epsilon}\right) t$.
\end{theorem}

@@

>>>>

Our main upper bound for polyhedral losses, Theorem \ref{thm:main-upper}, is mostly nonconstructive: it says polyhedral surrogates have some $\alpha$ such that $R_{\ell}(\psi \circ h; \D) \leq \alpha \cdot R_L(h;\D)$.
In this section, we make the result constructive by unpacking the implications of consistency (or more precisely, calibration) for polyhedral surrogates.

We will give a bound on the constant of the form $\alpha \leq \alpha_{\ell} \cdot \alpha_L \cdot \alpha_{\psi}$, i.e. one component each relating to the structure of the target loss, surrogate loss, and link.
First, we will derive $\alpha_L$ using the theory of \emph{Hoffman constants} from linear optimization.
Then, we will show that \emph{any} link for a consistent polyhedral loss must satisfy a condition called $\epsilon$-separation, introduced in \cite{finocchiario2020}.
We will set $\alpha_{\psi} = \frac{1}{\epsilon}$.
With $\alpha_{\ell}$ simply an upper bound on the target loss, we will finally prove $\alpha \leq \alpha_{\ell} \cdot \alpha_L \cdot \alpha_{\psi}$.

@@

Recall from the previous subsection that Hoffman constants allowed us to show that polyhedral expected losses grow linearly as we move away from the optimal set.
Now, $\epsilon$-separated link functions allow us to guarantee that we move at least a constant distance from the optimal set before we start linking to an ``incorrect'' report $r' \in \R$.
But when does a polyhedral loss have an associated $\epsilon$-separated link?

\paragraph{Existence of separated links.}
\cite{finocchiaro2020} showed that in fact, if $L$ indirectly elicits $\ell$, then there always exists an $\epsilon$-separated link $\psi$ such that $(L,\psi)$ is consistent for $L$.
Here, we strengthen this result much further, and give a converse that we believe is of independent interest: if $L$ indirectly elicits $\ell$, then \emph{every} link making $L$ consistent is $\epsilon$-separated for some $\epsilon > 0$.
Then, we show that this separation implies linear regret transfers (with a concrete constant).

\begin{lemma}\label{lemma:calibrated-eps-sep}
  Let polyhedral surrogate $L:\reals^d \to \reals^\Y_+$, discrete loss $\ell:\R\to\reals^\Y_+$, and link $\psi:\reals^d\to\R$ be given such that $(L,\psi)$ is calibrated with respect to $\ell$.
  Then there exists $\epsilon>0$ such that $\psi$ is $\epsilon$-separated with respect to   $\Gamma \defeq \prop{L}$ and $\gamma \defeq \prop{\ell}$.
\end{lemma}
\begin{proof}
  Suppose that $\psi$ is not $\epsilon$-separated for any $\epsilon>0$.
  Then letting $\epsilon_i \defeq 1/i$ we have sequences $\{p_i\}_i \subset \simplex$ and  $\{u_i\}_i \subset \reals^d$ such that for all $i\in\mathbb N$ we have both $\psi(u_i) \notin \gamma(p_i)$ and $d_\infty(u_i,\Gamma(p_i)) \leq \epsilon_i$.
  First, observe that there are only finitely many values for $\gamma(p_i)$ and $\Gamma(p_i)$, as $\R$ is finite and $L$ is polyhedral.
  \raft{Using a result from the polyhedral paper here, that $\Gamma$ only takes on finitely many values}
  Thus, there must be some $p\in\simplex$ and some infinite subsequence indexed by $j\in J \subseteq \mathbb N$ where
  for all $j\in J$, we have $\psi(u_j) \notin \gamma(p)$ and $\Gamma(p_j) = \Gamma(p)$.

@@

\subsection{Combining the loss and link}
We can now follow the general proof strategy of the main upper bound, but constructively.
Given Lemma \ref{lemma:lem:calibrated-eps-sep}, we can make the following definition.
\begin{definition}[$\epsilon_{\psi}$]
  If $(L,\psi)$ are calibrated for $\ell$, then let $\epsilon_{\psi} = \sup \{ \epsilon ~:~ \text{$\psi$ is $\epsilon$-separated}\}$.
\end{definition}
We will also need an upper bound on the regret of $\ell$.
\begin{definition}[$C_{\ell}$]
  Given a discrete target loss $\ell: \R \to \reals_+^{\Y}$, define $C_{\ell} = \max_{r,r' \in \R, y \in \Y} \ell(r)_y - \ell(r')_y$.
\end{definition}
Observe that $C_{\ell} = \max_{r,p} R_{\ell}(r,p)$.
\bo{needs justification?}
\bo{We could have $C_{\ell} = 0$ but then everything is optimal and the regret transfer is zero.}

@@

\begin{theorem}[Constructive linear transfer] \label{thm:separated-constant}
  Let $\ell: \R \to \reals_+^{\Y}$ be a discrete target loss, $L: \reals^d \to \reals_+^{\Y}$ be a polyhedral surrogate loss, and $\psi: \reals^d \to \R$ a link function.
  If $(L,\psi)$ are consistent for $\ell$, then
    \[ (\forall h,\D) \quad R_{\ell}(\psi \circ h ; \D) \leq \frac{C_{\ell} H_L}{\epsilon_{\psi}} R_L(h ; \D) ~. \]
\end{theorem}
The proof closely mirrors the proof of the nonconstructive upper bound, Theorem \ref{thm:main-upper}.
\begin{proof}
  By Lemma \ref{lemma:calibrated-eps-sep}, $\psi$ is separated and $\epsilon_{\psi}$ well-defined.
  By Lemma \ref{lemma:separated-constant-p}, for each $p \in \mathcal{Q}$, $R_{\ell}(\psi(u),p) \leq \frac{C_{\ell} H_L}{\epsilon_{\psi}} R_L(u,p)$ for all $u$.
  Now consider a general $p$, which is in some full-dimensional polytope level set $\Gamma_u$.
  Write $p = \sum_{q \in \mathcal{Q}_u} \beta(q) q$ for some probability distribution $\beta$, where $\mathcal{Q}_u$ is the set of vertices of $\Gamma_u$.
  By Lemma \ref{lemma:linear-on-levelset}, $R_L$ and $R_{\ell}$ are linear in $p$ on $\Gamma_u$, so for any $u'$,
  \begin{align*}
    R_{\ell}(\psi(u'),p)
    &=    \sum_{q \in \mathcal{Q}_u} \beta(q) R_{\ell}(\psi(u'), q)  \\
    &\leq \sum_{q \in \mathcal{Q}_u} \beta(q) \frac{C_{\ell} H_{L,p}}{\epsilon_{\psi}} R_L(u', q)  \\
    &\leq \frac{C_{\ell} H_L}{\epsilon_{\psi}} \sum_{q \in \mathcal{Q}_u} \beta(q) R_L(u', q)  \\
    &\leq \frac{C_{\ell} H_L}{\epsilon_{\psi}} R_L(u', p) .
  \end{align*}
  This conditional regret transfer implies a full regret transfer with the same constant by Observation \ref{obs:transfer}.
\end{proof}

@@

\begin{theorem} \label{thm:tfae}
  Let $\ell: \R \to \reals_+^{\Y}$ be a discrete target loss, $L: \reals^d \to \reals_+^{\Y}$ be a polyhedral surrogate loss, and $\psi: \reals^d \to \R$ a link function.
  The following are equivalent:
  \begin{enumerate}
    \item $(L,\psi)$ are consistent for $\ell$.
    \item $(L,\psi)$ indirectly elicit $\ell$ and $\psi$ is $\epsilon$-separated for some $\epsilon > 0$.
    \item $(L,\psi)$ guarantee a linear regret transfer for $\ell$.
  \end{enumerate}
  Furthermore, if any of the above hold, then in particular $(L,\psi)$ guarantee the regret transfer $\zeta(t) = \left(\frac{C_{\ell} H_L}{\epsilon_{\psi}}\right) t$.
\end{theorem}

@@

\bo{Observe that our theory gives some useful way to compare link functions that may not have existed before, i.e. bigger $\epsilon$ seems to be better. Full stop?}
